Item Construction Rules Revisited: Learnings from Measurement of Latent Variables with Gold Standard Items

### 2025-06-17 manuscript.qmd
tbl-3 note
add correlations to gold standard item discussion

```{r}
#| label: "tbl-3"
#| tbl-cap: Standardized item loadings on the latent variable for two and six categories controlling for the self-reported physical item and self-reported gender
#| ft.align: left
#| disable-apaquarto-processing: false
#| apa-note: Standardized regression weights of the self-reported physical item and self-reported gender greater than .60 are printed in bold typeface. LV = TODO

@@

Gold-standard items did not work as theoretically but empirically expected in this study. However, we still believe they are important. Only the gold standard item for body height – for six, not for two response categories – had its expected highest loading on the latent variable. @bortolottiRelevanceAdvantagesUsing2013, who previously applied the same procedure with two response categories to investigate a gold-standard item for body height, found a correlation of .86 with body height. This correlation pretty much matches our correlation with the latent variable. For our 2-response category scale, the correlation was slightly higher (.87), but there were also items which had descriptively equal or higher correlations (biserial correlations) with the sum score than the self-reported physical item (Pearson correlation). This might be because we reported a latent correlation corrected for diattenuation, and our scale had 12 instead of 27 items. Nevertheless, our results depict a conceptual replication of @bortolottiRelevanceAdvantagesUsing2013 with a different questionnaire.

>>>>

```{r}
#| label: "tbl-3"
#| tbl-cap: Standardized item loadings on the latent variable for two and six categories controlling for the self-reported physical item and self-reported gender
#| ft.align: left
#| disable-apaquarto-processing: false
#| apa-note: Standardized regression weights of the self-reported physical item and self-reported gender greater than .60 are printed in bold typeface. LV = latent variable (height, weight, age).

@@

Gold-standard items did not work as theoretically but empirically expected in this study. However, we still believe they are important. Only the gold standard item for body height – for six, not for two response categories – had its expected highest loading on the latent variable. @bortolottiRelevanceAdvantagesUsing2013, who previously applied the same procedure with two response categories to investigate a gold-standard item for body height, found a correlation of .86 with body height. This correlation pretty much matches our correlation with the latent variable. The correlation was slightly higher for our 2-response category scale (`r results$load_df_nomod |> filter(rhs == "Koerpergroesse_realW" & categories == 2) |> pull(est.std) |> printnum(gt1 = FALSE)`) compared to our 6-response category scale (`r results$load_df_nomod |> filter(rhs == "Koerpergroesse_realW" & categories == 6) |> pull(est.std) |> printnum(gt1 = FALSE)`), but there were also items which had descriptively equal or higher correlations (biserial correlations) with the sum score than the self-reported physical item (Pearson correlation). This might be because we reported a latent correlation corrected for disattenuation, and our scale had 12 instead of 27 items. Nevertheless, our results depict a conceptual replication of @bortolottiRelevanceAdvantagesUsing2013 with a different questionnaire.
